Review: Crop adaptation through introgression from wild relatives

### 2017-04-11 paper.tex
Update paper.tex

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{authblk}%allows footnote format for authors
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\usepackage[numbers,sort]{natbib}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{adjustbox}
\bibliographystyle{ieeetr}
\newcommand{\mbh}[1]{\textcolor{orange}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
\newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments
\newcommand{\gmj}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments

@@

Whereas the process of domestication used to be concieved of as rapid and geographically-constrained, archaeological and genetic evidence has recently revealed that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
This new paradigm emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.
Adaptive introgression has two steps: hybridization between two genomes (in the case of this review, a domesticated crop and its wild relative), followed by recombination and selection for progressively smaller genomic segments that harbor the adaptive locus/loci with fewer maladaptive alleles being dragged along \cite{Feuillet200824}.

The traditional paradigm in crop domestication has been origin from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
Absent from this paradigm are the effects of hybridization between a diffusing nascent crop species and closely-related, locally-adapted wild relatives outside the center of origin.
New methods have recently been employed to detect genome-wide patterns of introgression in a number of species.
In this review, we will: 1) briefly describe these methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.

@@

The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
Pre-Columbian \emph{H. annuus} distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H. petiolaris} and \emph{H. bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
These introgressed hybrids formed a new lineage of sunflower (\emph{H. annuus} ssp. \emph{texanus}, \emph{H. a. texanus} hereafter) which displays \emph{H. debilis}-like traits adaptive to south-central Texas climate and ecology.
These adaptive debilis-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H. annuus} (as measured by higher seed production)\cite{whitney2006adaptive}.
Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.

@@

\begin{table}
\centering
\begin{adjustbox}{width=1\textwidth}
\small
\label{my-label}
\begin{tabular}{llllll}
\hline
Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Introgression into Crop & Evidence of Adaptiveness & Source \\
\hline
Maize (Zea mays subsp. mays) & Zea mays subsp. mexicana, Zea mays subsp. parviglumis & X & X & X & \cite{hufford2013genomic} \\
Rice (Oryza sativa, Oryza Glaberrima) & Oryza rufipogon & X & X & X & \cite{Huang2012} \\
Barley (Hordeum vulgare) & Hordeum vulgare subsp. spontaneum & X & X & X & \cite{Poets2015} \\
Cassava (Manihot esculenta) & Manihot glaziovii & X & X & X & \cite{bredeson2016sequencing} \\
Olive (Olea europaea ssp. europaea var. sativa) & [[]] & X & X & & \cite{diez2015olive} \\
Sunflower (Helianthus annuus) & wild H. annuus, H. argophyllus, H. petiolaris & X & X & X & [[]]\\
Potato (Solanum tuberosum) & many & X & X & X & \cite{johns1986ongoing} \\
Soybeans (Glycine max) & G. soja & [[]] & X & X & \cite{lam2010resequencing} \\
Grapes (Vitis vinifera subsp. vinifera) & Vitis vinifera subsp. sylvestris & [[]] & X & [[]] &  \cite{myles2011genetic} \\
\hline
\end{tabular}
\end{adjustbox}
\end{table}

@@

\section*{Re-evaluating concepts of domestication}
A framework in which crops were domesticated from a single population or even a single species is, in several instances, an oversimplification. A demography incorporating introgression from additional sources appears to be more correct for many crops. With this in mind, certain aspects of crop evolution must be re-evaluated:
\begin{enumerate}
	\item{Estimates of the initial domestication bottleneck may be skewed when introgression is not considered. Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.  For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.}%\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?  I just want to make sure I understand.}
	\item{Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.}%\lwang{It could be either inflated or underestimate,. It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}
	\item{Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa. Introgression may alter these signatures and confound detection of domestication loci.} %\lwang{I donot quite understand this point. So, it is assumed that the introgressed haplotype has higher diversity, which renders difficulty to detect selection in such regions, right? But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
	%\gmj{It may not fit here perfectly, but you might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.  The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}
\end{enumerate}

\section*{Future studies in crop-wild introgression}
Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics:
\begin{enumerate}
	\item{Basic questions:}
		\begin{enumerate}
		 	\item{To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa?} %\lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
		 	\item{At what geographic scale does adaptive introgression occur? Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges? To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?}%\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g. temperature, precipitation, altitude, day lengh, etc.)?}
		 	\item{Can colonizing species served as bridges for gene flow between previously allopatric taxa?} %\lwang{I feel the second and the third points could be combined. How the wild species could introgress into the allopatric landraces? Does domesticates serve as bridges?}
		\end{enumerate} 	
	\item{Applications:}
		\begin{enumerate}
		 	\item{Loci underlying the domesticated phenotype which may be beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.} %\lwang{I am not quite clear with this point. It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow. Right?}\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?  I hope that question made sense.}
		 	\item{Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.}%\gmj{This point is very clear, but should it be fleshed out further?  It wouldn't be difficult to include some examples.}
		 \end{enumerate}
\end{enumerate}

>>>>

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{authblk}%allows footnote format for authors
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\usepackage[numbers,sort]{natbib}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage[english]{babel}
\bibliographystyle{ieeetr}
\newcommand{\mbh}[1]{\textcolor{orange}{ \emph{\scriptsize  #1}} } %creating command for Matt's comments
\newcommand{\lwang}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %creating command for Li's comments
\newcommand{\gmj}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %creating command for Garrett's comments

@@

The process of domestication was one thought to be rapid and geographically-constrained, with origin from a wild relative within one or more geographically-defined centers followed by expansion to the modern-day extent of cultivation.
However, archaeological and genetic evidence are revealing that domestication is more often temporally-protracted and multiregional \cite{brown2009complex}.
This new conception of domestication emphasizes the role of gene flow of beneficial alleles into domesticating populations from sympatric, locally-adapted wild relatives, a process known as adaptive introgression.

Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
In the case of adaptive introgression into domesticated crops, crop/wild hybrids backcross to the domesticate and adaptive genomic regions are retained and increase in frequency in the population as recombination separates desirable alleles from neighboring maladaptive alleles, which are selected against and lost.
Much attention has been given to the risk of natural introgression of transgenes from domessticated crops into wild relatives (for a review, \cite{stewart2003transgene}), or the history, utility, and methods of breeding programs designed to introgress desired traits from wild relatives (for a review, \cite{zamir2001improving, tanksley1997seed, hajjar2007use}), but recent tools and methods have been employed to detect genome-wide patterns of introgression in a number of species, granting insights into the prominence of adaptive introgression in crop domestication.

In this review, we will: 1) briefly describe these methods and provide a summary of their application for detecting crop-wild introgression, 2) review evidence supporting the hypothesis that wild-to-crop introgression has conferred local adaptation, 3) consider how the prevalence of this introgression alters traditional concepts of domestication, and 4) describe future advances in both basic and applied genetics that can be made through the study of introgression in agroecosystems.

@@

The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
Pre-Columbian \emph{H. annuus} distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H. petiolaris} and \emph{H. bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
These introgressed hybrids formed a new lineage of sunflower (\emph{H. annuus} ssp. \emph{texanus}, \emph{H. a. texanus} hereafter) which displays \emph{H. debilis}-like traits adaptive to south-central Texas climate and ecology.
These adaptive \emph{debilis}-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H. annuus} (as measured by higher seed production)\cite{whitney2006adaptive}.
Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.

@@

\begin{table}
\centering
\begin{adjustbox}{width=1\textwidth}
\small
\label{my-label}
\begin{tabular}{|p{5cm}|p{5cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}|l|}
\hline
Crop & Compatible Wild Relatives & Hybrids and/or Hybridization & Evidence of Crop Introgression & Evidence of Adaptiveness & Source \\ \hline \hline
Maize (\emph{Zea mays} subsp. \emph{mays}) & \emph{Z. m.} subsp. \emph{mexicana}, \emph{Z. m. } subsp. \emph{parviglumis} & X & X & X & \cite{hufford2013genomic} \\ 
\hline 
Rice (\emph{Oryza sativa}) & \emph{O. rufipogon} & X & X & X & \cite{Huang2012} \\ 
\hline
Barley (\emph{Hordeum vulgare}) & \emph{H. v.} subsp. \emph{spontaneum} & X & X & X & \cite{Poets2015} \\ \hline
Sunflower (\emph{Helianthus annuus}) & \emph{H. argophyllus}, \emph{H. bolanderi}, \emph{H. debilis}, \emph{H. petiolaris} & X & X & X & \cite{rieseberg2007hybridization}\\ 
\hline
Cassava (\emph{Manihot esculenta}) & \emph{M. glaziovii} & X & X & X & \cite{bredeson2016sequencing} \\ 
\hline
Potato (\emph{Solanum tuberosum}) & many & X & X & X & \cite{johns1986ongoing} \\ \hline
Olive (\emph{Olea europaea} ssp. \emph{europaea} var. \emph{sativa}) & \emph{O. e.} ssp. \emph{europaea} var. \emph{sylvestris} & X & X & & \cite{diez2015olive} \\ 
\hline
Soybeans (\emph{Glycine max}) & \emph{G. soja} & X & X &  & \cite{lam2010resequencing} \\ 
\hline
Grapes (\emph{Vitis vinifera} subsp. \emph{vinifera}) & \emph{V. v.} subsp. \emph{sylvestris} & X & X &  &  \cite{myles2011genetic} \\
\hline
\end{tabular}
\end{adjustbox}
\end{table}

@@

\section*{Re-evaluating concepts of domestication}
A framework in which crops were domesticated from a single population or even a single species is, in several instances, an oversimplification.
A demography incorporating introgression from additional sources appears to be more correct for many crops.
With this in mind, certain aspects of crop evolution must be re-evaluated:

Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.

%\lwang{if the donor population has a relatively small population size, then the domestication bottleneck could be over-estimated.}\gmj{is the Kelley Harris, Rasmus Nielsen paper on Neanderthal genetic load in humans (the old CAMP paper) an example of this, Li?  I just want to make sure I understand.}
Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
%\lwang{It could be either inflated or underestimate,. It depends on the introgressed haplotype are more or less divergent from the wild progenitor, right?}

Loci under selection during domestication are often identified based on signatures of substantially-reduced nucleotide diversity in the domesticated taxon relative to the wild progenitor and high allele frequency differentiation between these taxa.
Introgression may alter these signatures and confound detection of domestication loci.
%\lwang{I donot quite understand this point. So, it is assumed that the introgressed haplotype has higher diversity, which renders difficulty to detect selection in such regions, right? But the case in maize is that even the introgressed haplotypes had higher diversity, but as selection has act long time on the region, the observed current diversity has already been reduced.}
%\gmj{It may not fit here perfectly, but you might consider a point about how crop-wild introgressions can also mask or obscure even the true progenitor species and center(s) of domestication.  The examples of potato and tomato domestication comes to mind; in potato, persistent and thorough interbreeding with a complex of wild relatives during domestication probably make it impossible to identify a single progenitor (if ever there was one), and in tomato, widescale recent introgression between two or three wild relatives confound attempts to identify which is the progenitor, and therefore the center of domestication as well.}

\section*{Future studies in crop-wild introgression}

New insights into domestication histories of crops are of potential value for the continued improvement of modern crops.
Continued research into this crop/wild introgression will prove valuable

Additional study of introgression in agroecosystems could lead to advances in both basic and applied genetics.
Basic questions:
To what extent does the level of introgression across taxa depend on divergence time between donor and recipient taxa? %\lwang{maybe put alternative hypothesis here, or due to relative mutation load of the donor and recipient species?}
At what geographic scale does adaptive introgression occur? Is introgression frequently restricted to very local populations or is it often seen over broad geographic ranges? To what extent does this depend on the slope of environmental gradients such as temperature, precipitation, and elevation?%\gmj{Maybe not just geographical ranges, but also across environmental clines (e.g. temperature, precipitation, altitude, day lengh, etc.)?}
Can colonizing species served as bridges for gene flow between previously allopatric taxa? %\lwang{I feel the second and the third points could be combined. How the wild species could introgress into the allopatric landraces? Does domesticates serve as bridges?

Loci underlying the domesticated phenotype which may be beneficial targets for crop improvement can be more clearly identified by removing the confounding population genetic signal of introgression.
%\lwang{I am not quite clear with this point. It will be interesting to see that the domestication loci have further receive gene flow from the other wild relatives and it has not been purged out, as I think the domestication loci undergo strong positive selection, thus also strong purifying selection to dispel influx gene flow. Right?}
%\gmj{Removing the signal of introgression, is this more likely to uncover new and interesting domestication loci, or will it more likely reveal that some of the previously-identified loci of interest are not very interesting after all (as in, the identification of the introgression changes our understanding of the history of the locus and there is less evidence of selective pressures on it), or equally either case?  I hope that question made sense.}
Adaptive introgression that is clearly tied to a specific environment may include beneficial alleles that can be utilized in crop breeding.
